SciPy 1.0---Fundamental Algorithms for Scientific Computing in Python

### 2018-07-25 scipy-1.0/paper.tex
Drafted two short paragraphs for sparse tech improvements section.

\subsection*{API and ABI evolution}
The application programming interface (API) for SciPy consists of
approximately 2600 functions and classes. Adding new functionality has to
meet the following criteria:
\begin{itemize}
    \item The algorithm is of relevance to multiple fields of science.
    \item The implementation is demonstrably important.  E.g., it is classic
    enough to be included in textbooks, or it is based on a peer-reviewed article
    which has a significant number of citations.
\end{itemize}

Our policy for evolving the API over time is that new functionality can be added,
while removing or changing existing functionality can only be done if the benefits
of that exceeds the (often significant) costs to users, \textit{and} only
after giving clear deprecation warnings to those users for at least one year.
In general, we encourage changes that improve clarity in the API of the
library but strongly discourage breaking backwards compatibility\footnote{
    Backwards compatibility means that new versions of SciPy will keep user
    code written for an older version working.}
given our position near the base of the scientific Python computing stack.

@@

@@

\subsection*{Statistical distributions}

>>>>

\subsection*{API and ABI evolution}
The application programming interface (API) for SciPy consists of
approximately 1500 functions and classes.
Our policy for evolving the API over time is that new functionality can be added,
while removing or changing existing functionality can only be done if the benefits
of that exceeds the (often significant) costs to users, \textit{and} only
after giving clear deprecation warnings to those users for at least one year.
In general, we encourage changes that improve clarity in the API of the
library but strongly discourage breaking backwards compatibility\footnote{
    Backwards compatibility means that new versions of SciPy will keep user
    code written for an older version working.}
given our position near the base of the scientific Python computing stack.

@@

Over the last three years our sparse matrix handling internals have
been rewritten and performance has been improved. Iterating over 
and slicing of CSC and CSR matrices is now faster by up to 35\%,
and the CSR to COO matrix format conversion is now several times faster.
COO matrices are now generally more memory efficient. Importantly,
SuperLU\cite{superlu_ug99} was updated to version 5.2.1, enhancing
the low-level implementations leveraged by a subset of our \texttt{sparse}
offerings.

From a new features standpoint, \texttt{scipy.sparse} matrices and linear
operators now support the Python matrix multiplication (@) operator when
available. We've added \texttt{scipy.sparse.norm} and \texttt{scipy.sparse.random} 
for computing sparse matrix norms and drawing random variates from arbitrary 
distributions, respectively. Users can now easily serialize a number of our
sprase matrix formats with the newly associated \texttt{save\_npz} function.
We've enhanced the diagonal selection API for sparse matrices to allow
for more user control.

@@

\subsection*{Statistical distributions}
The \texttt{scipy.stats} module provides provides more than 100 probability
distributions and a framework to implement additional distributions. Given
an expression for a probability distribution function (pdf),
the framework will calculate moments, cumulative distribution function (cdf),
differential entropy, and more. Those properties are calculated based on
generic code, e.g., numerical integration of the pdf to obtain the cdf.
Where possible, the generic code is replaced by explicit formulas or more
accurate or efficient implementations. For example, most distributions have
an explicit formula for the cdf, so it is used instead of integrating the
pdf. The distributions framework also allows the user to draw random
variates from the distribution. Drawing random variates now accepts a \texttt{random\_state}
parameter, which is either a \texttt{numpy.random.RandomState} object or an
integer used to generate such an object, to provide fine-grained user-level
control over the random state of a workflow.
Currently about 80 continuous univariate distributions, 12 discrete univariate
distributions and 10 multivariate distributions are implemented.
Key recent distributions added to \texttt{scipy.stats} include the histogram-based
distribution in \texttt{scipy.stats.rv\_histogram} and the multinomial
distribution in \texttt{scipy.stats.multinomial} (used, for example, in natural
language processing\cite{Griffiths5228}). \texttt{multinomial} is the first and so 
far the only multivariate discrete distribution available in SciPy.
